Identifying Inter-Genomic Repeats Using Fast, Approximate Methods for Betweenness Centrality

### 2015-05-19 manuscript/manuscript.tex
Edited methods.

We evaluate the efficacy of detecting inter-genomic repeats using the approximate betweenness centrality methods on simulated metagenomic assembly graphs.
The first step is to construct the \textit{de Bruijn} graph from the set of reference genomes.
De Bruijn strings are a type of ``comprehensive'' strings. For a given alphabet $\Sigma = \{A,C,T,G\}$ and a length $k$, a de Bruijn string of order $k$ contains as substrings of length $k$ over $\Sigma$. These strings can be represented in a sophisticated way in term of a graph called de Bruijn graph \cite{debruijn}. A de Bruijn graph of order $k$ is a graph that contains all strings of length $k-1$ from a given alphabets as nodes and contains edges between two nodes if the corresponding strings overlap by exactly $k-2$ letters in prefix-suffix manner. In other words, each $k$-mer is represented by two nodes in a graph connected by an edge.

Next we modify the de Bruijn graph generated from genomes to identify repeats. We define \textit{unipaths} as the path of nodes in which the outdegree of first node on the path is $1$ and indegree is not equal to $1$, indegree of last node on the path is $1$ and outdegree is not equal to $1$ and for all other nodes along the path, both indegree and out degree is exactly $1$. We compress all the unipaths in a graph to represent the path in a compact way. We represent the unipath as a single node whose label is the concatenation of all node labels along the path.
Nodes in the resulting graph represent uniquely resolvable sequences of length $\geq k$.

@@

In addition to the simulated metagenomic assembly graphs, we study how approximation algorithms for betweenness centrality perform on real datasets. We evaluated the runtime  of these algorithms on a metagenomic stool sample of a premature infant\cite{morowitz2011strain}. This graph has 3,797,425 nodes and 2,970,379 edges. We compared the runtime of Bambus2's repeat detection module on this data with the runtime of approximate betweenness centrality algorithms.

>>>>

We evaluate the efficacy of detecting inter-genomic repeats using the approximate betweenness centrality methods on simulated metagenomic assembly graphs.
The first step is to construct the \textit{de Bruijn} graph from the set of reference genomes.
De Bruijn strings are a type of ``comprehensive'' strings. For a given alphabet $\Sigma = \{A,C,T,G\}$ and a length $k$, a de Bruijn string of order $k$ contains as substrings of length $k$ over $\Sigma$. These strings can be represented in a sophisticated way in term of a graph called de Bruijn graph \cite{debruijn}. A de Bruijn graph of order $k$ is a directed graph that contains all strings of length $k-1$ from a given alphabets as nodes and contains edges between two nodes if the corresponding strings overlap by exactly $k-2$ letters in prefix-suffix manner. In other words, each $k$-mer is represented by two nodes in a graph connected by an edge.

Next we modify the de Bruijn graph generated from the collection of genomes to identify repeats $>=k$. We define \textit{unipaths} as the path of nodes in which the outdegree of first node on the path is $1$ and indegree is not equal to $1$, indegree of last node on the path is $1$ and outdegree is not equal to $1$ and for all other nodes along the path, both indegree and out degree is exactly $1$. We compress all the unipaths in a graph to represent the path in a compact way. We represent the unipath as a single node whose label is the concatenation of all node labels along the path.
Nodes in the resulting graph represent uniquely resolvable sequences of length $\geq k$.

@@

In addition to the simulated metagenomic assembly graphs, we evaluate the runtimes of the two approximation algorithms for betweenness centrality on a metagenomic stool sample of a premature infant\cite{morowitz2011strain}. This graph has 3,797,425 nodes and 2,970,379 edges. We compared the runtime of Bambus2's repeat detection module on this data with the runtime of approximate betweenness centrality algorithms.
